---
output: github_document
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      include = TRUE)
```


```{r}
source("2-code/0-packages.R")
targets::tar_load_everything()
```

## FTICR-MS

### VAN KREVELEN

```{r}
gg_vankrev <- function(data,mapping){
  ggplot(data,mapping) +
    # plot points
    geom_point(size=0.5, alpha = 0.5) + # set size and transparency
    # axis labels
    ylab("H/C") +
    xlab("O/C") +
    # axis limits
    xlim(0,1.25) +
    ylim(0,2.5) +
    # add boundary lines for Van Krevelen regions
    geom_segment(x = 0.0, y = 1.5, xend = 1.2, yend = 1.5,color="black",linetype="longdash") +
    geom_segment(x = 0.0, y = 0.7, xend = 1.2, yend = 0.4,color="black",linetype="longdash") +
    geom_segment(x = 0.0, y = 1.06, xend = 1.2, yend = 0.51,color="black",linetype="longdash") +
    guides(colour = guide_legend(override.aes = list(alpha=1, size = 1)))
}

```

Unique peak assignments for + vs. - mode
```{r}
# vk_domains = 
lcms_metadata %>% 
  gg_vankrev(aes(x = OC, y = HC, color = mode))+
  theme_kp()+
  facet_wrap(~mode)


print("unique + vs - mode")
lcms_metadata %>% 
  filter(mode == "negative, positive") %>% 
  ggplot(aes(x = OC, y = HC))+
  geom_point(data = lcms_metadata %>% filter(mode != "negative, positive"),
             aes(color = mode))+
  #   geom_point(color = "grey", alpha = 0.8, size = 1) + # set size and transparency
  # axis labels
  ylab("H/C") +
  xlab("O/C") +
  # axis limits
  xlim(0,1.25) +
  ylim(0,2.5) +
  # add boundary lines for Van Krevelen regions
  geom_segment(x = 0.0, y = 1.5, xend = 1.2, yend = 1.5,color="black",linetype="longdash") +
  geom_segment(x = 0.0, y = 0.7, xend = 1.2, yend = 0.4,color="black",linetype="longdash") +
  geom_segment(x = 0.0, y = 1.06, xend = 1.2, yend = 0.51,color="black",linetype="longdash") +
  guides(colour = guide_legend(override.aes = list(alpha=1, size = 2)))+
  scale_color_manual(values = whistledown::whistledown_palette("queen", 3))


```

All identified peaks across the different samples

```{r lcms_vk, fig.width=12, fig.height=9}

lcms_long_combined %>% 
  left_join(lcms_metadata) %>% 
  gg_vankrev(aes(x = OC, y = HC, color = mode))+
  facet_grid(horizon ~ site + transect_location)

```


### relabund

```{r}

lcms_relabund_presence_cores %>% 
  ggplot(aes(x = orbitrap_id, y = rel_abund, fill = Class_detailed)) +
    geom_bar(stat = "identity", position = "stack") +
    facet_wrap(~site, scales = "free_x", ncol = 6) +
    labs(x = "",
         y = "Relative abundance (%)")+
    scale_fill_brewer(palette = "RdYlBu") +
    theme_bw() +
    theme(legend.title = element_blank()) +
    theme(axis.text.x = element_text(angle = 90)) 

lcms_relabund_presence_treatment %>% 
  ggplot(aes(x = sample_type, y = mean, fill = Class_detailed)) +
    geom_bar(stat = "identity", position = "stack") +
    facet_wrap(~site + mode, scales = "free_x", ncol = 6) +
    labs(x = "",
         y = "Relative abundance (%)")+
    scale_fill_brewer(palette = "RdYlBu") +
    theme_bw() +
    theme(legend.title = element_blank()) +
    theme(axis.text.x = element_text(angle = 90)) 

lcms_relabund_intensities_cores %>% 
  ggplot(aes(x = orbitrap_id, y = rel_abund, fill = Class_detailed)) +
    geom_bar(stat = "identity", position = "stack") +
    facet_wrap(~site, scales = "free_x", ncol = 6) +
    labs(x = "",
         y = "Relative abundance (%)")+
    scale_fill_brewer(palette = "RdYlBu") +
    theme_bw() +
    theme(legend.title = element_blank()) +
    theme(axis.text.x = element_text(angle = 90)) 


lcms_relabund_intensities_treatment %>% 
  ggplot(aes(x = sample_type, y = mean, fill = Class_detailed)) +
    geom_bar(stat = "identity", position = "stack") +
    facet_wrap(~site + mode, scales = "free_x", ncol = 6) +
    labs(x = "",
         y = "Relative abundance (%)")+
    scale_fill_brewer(palette = "RdYlBu") +
    theme_bw() +
    theme(legend.title = element_blank()) +
    theme(axis.text.x = element_text(angle = 90)) 


```






### PERMANOVA

```{r}

adonis2(wide_pos_formula %>% dplyr::select(where(is.numeric)) ~ 
          (site + transect_location + horizon),
        by = "margin",
        data = wide_pos_formula) %>% 
  knitr::kable()

 adonis2(wide_pos_features %>% dplyr::select(where(is.numeric)) ~ 
           (site + transect_location + horizon),
         by = "margin",
         data = wide_pos_formula) %>% 
   knitr::kable()

adonis2(wide_neg_formula %>% dplyr::select(where(is.numeric)) ~ 
          (site + transect_location + horizon),
        by = "margin",
        data = wide_pos_formula) %>% 
  knitr::kable()

 adonis2(wide_neg_features %>% dplyr::select(where(is.numeric)) ~ 
           (site + transect_location + horizon),
         by = "margin",
         data = wide_pos_formula) %>% 
   knitr::kable()

```


### PCOA

```{r}

plot_lcms_pcoa = function(dat){
  
  pacman::p_load("BiocManager","ComplexHeatmap","dendextend","NbClust","cowplot",
                 "gtools", # mixedsort
                 "tidyverse",
                 "janitor", # row_to_names function
                 "factoextra", # PCA
                 "vegan") # PCoA
  
  
  df = dat %>% column_to_rownames("orbitrap_id") %>% dplyr::select(where(is.numeric))
  grp = dat %>%  dplyr::select(!where(is.numeric))
  
  
  # PCoA
  distm <- vegdist(df, method = "bray", na.rm = FALSE) #compute distance
  
  PcoA <- cmdscale(distm, k = 2, eig = T, add = T)
  
  PcoA_points <- as.data.frame(PcoA$points) #getting the PCOs into dataframe
  variance <- round(PcoA$eig*100/sum(PcoA$eig),1) # getting the variance explained by each PCo
  names(PcoA_points)[1:2] <- c("PCoA1", "PCoA2")
  
  PcoA_points2 = PcoA_points %>% rownames_to_column("orbitrap_id") %>% left_join(grp)
  
  ggplot(PcoA_points2, 
         aes(x = PCoA1, 
             y = PCoA2,
             colour = as.factor(sample_type), 
             label = orbitrap_id)) +
    geom_point(size=4, alpha = 1) +
  #  geom_text_repel() +
  #  scale_colour_manual(values = c('orange','darkgreen','red','blue','black','grey','purple', 'skyblue3','magenta','green')) +
    xlab(paste('PCoA1',variance[1],'%', sep = ' ')) + 
    ylab(paste('PCoA2',variance[2],'%', sep = ' ')) + 
    labs(color = '') +
    coord_fixed() +
    theme(axis.text = element_text(size = 16),
          axis.title = element_text(size = 16, face= 'bold'),
          plot.title = element_text(size = 18, face= 'bold',hjust=0.5),
          legend.title = element_text(size = 18, face= 'bold'),
          legend.text = element_text(size = 16),
          panel.background = element_blank(),
          panel.border = element_rect(colour = "black", fill=NA, linewidth=1))
  
  
}
plot_lcms_pcoa(dat = wide_pos_features)
plot_lcms_pcoa(dat = wide_neg_features)

```



















So we split the data by mode for all subsequent analyses.
```{r}
adonis2(lcms_relabundance_wide %>% filter(mode == "negative") %>% 
          dplyr::select(where(is.numeric)) ~ 
          (region + transect_location + horizon),
        by = "margin",
        data = lcms_relabundance_wide %>% filter(mode == "negative")) %>% 
  knitr::kable()
```
For negative mode, region accounted for 66 % and transect 25 % of total variability.


```{r}
adonis2(lcms_relabundance_wide %>% filter(mode == "positive") %>% 
          dplyr::select(where(is.numeric)) ~ 
          (region + transect_location + horizon),
        by = "margin",
        data = lcms_relabundance_wide %>% filter(mode == "positive")) %>% 
  knitr::kable()
```
For positive mode, region accounted for 34 % and transect 23 % of total variability.



### PCA

```{r pca_function}
fit_pca_function_lcms = function(dat, sample_key){
  
  relabund_pca =
    dat %>% 
    replace(.,is.na(.),0)
  
  num = 
    relabund_pca %>% 
    dplyr::select(where(is.numeric))
  
  grp = 
    relabund_pca %>% 
    dplyr::select(where(is.character)) %>% 
    dplyr::mutate(row = row_number())
  
  pca_int = prcomp(num, scale. = T)
  
  list(num = num,
       grp = grp,
       pca_int = pca_int)
  
}
```

```{r}
lcms_pca_positive = fit_pca_function_lcms(dat = lcms_relabundance_wide %>% filter(mode == "positive"))
lcms_pca_negative = fit_pca_function_lcms(dat = lcms_relabundance_wide %>% filter(mode == "negative"))
```

```{r lcms_pca_biplot}
ggbiplot(lcms_pca_positive$pca_int, obs.scale = 1, var.scale = 1,
         groups = lcms_pca_positive$grp$region, 
         ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(shape = lcms_pca_positive$grp$transect_location,
                 color = groups))+
  scale_shape_manual(breaks = c("upland", "transition", "wte", "wetland"),
                     values = c(1,2,3,4))+
  xlim(-4,4)+
  ylim(-3.5,3.5)+
  labs(color = "", shape = "",
       title = "FTICR - PCA - positive mode")+
  #   theme_kp()+
  theme(legend.position = "top", legend.box = "vertical")+
  NULL


ggbiplot(lcms_pca_negative$pca_int, obs.scale = 1, var.scale = 1,
         groups = lcms_pca_negative$grp$region, 
         ellipse = TRUE, circle = FALSE, var.axes = TRUE, alpha = 0) +
  geom_point(size=3,stroke=1, alpha = 1,
             aes(shape = lcms_pca_negative$grp$transect_location,
                 color = groups))+
  scale_shape_manual(breaks = c("upland", "transition", "wte", "wetland"),
                     values = c(1,2,3,4))+
  xlim(-4,4)+
  ylim(-3.5,3.5)+
  labs(color = "", shape = "",
       title = "FTICR - PCA - negative mode")+
  #   theme_kp()+
  theme(legend.position = "top", legend.box = "vertical")+
  NULL


```

### NMDS

```{r lcms_cluster_formula}
library(cluster)
library(factoextra)

lcms_wide = 
  lcms_data_trt %>% 
  distinct(formula, sample_name, .keep_all = T) %>% 
  dplyr::select(formula, intensity, sample_name, sample_type, region, site, transect_location, horizon, mode) %>% 
  pivot_wider(names_from = "formula", values_from = "intensity") %>% 
  janitor::remove_empty("cols") %>% 
  replace(is.na(.), 0)

df = fticr_wide %>%  dplyr::select(where(is.numeric))
grp = fticr_wide %>%  dplyr::select(!where(is.numeric))


```


clustering and NMDS for NEGATIVE MODE

```{r, include=FALSE}


compute_clustering = function(dat){
  
  
  df = dat %>%  dplyr::select(where(is.numeric))
  grp = dat %>% dplyr::select(!where(is.numeric))
  
  
  fviz_nbclust(df, FUN = hcut, method = "wss")
  fviz_nbclust(df, FUN = hcut, method = "silhouette")
  # two clusters
  
  d <- dist(df, method = "binary")
  hc1 <- hclust(d, method = "complete" )
  plot(hc1, cex = 0.6, hang = -1)
  hc5 <- hclust(d, method = "ward.D2" )
  sub_grp <- cutree(hc5, k = 3)
  # table(sub_grp)
  plot(hc5, cex = 0.6)
  # rect.hclust(hc5, k = 5, border = 2:5)
  
  df_cl <- mutate(df, cluster = sub_grp) %>% cbind(grp)
  
  cluster = df_cl %>% dplyr::select(sample_type, cluster) %>% 
    dplyr::mutate(row = row_number())
  
  
  # print("hierarchical clustering: three clusters")
  # count(df_cl,cluster)
  
}

clusters_neg = compute_clustering(dat = fticr_wide %>% filter(mode == "negative"))
clusters_pos = compute_clustering(dat = fticr_wide %>% filter(mode == "positive"))


```




```{r icr_mds_calc, include=FALSE}

compute_nmds = function(dat, cluster){
  
  num = 
    dat %>% 
    #  column_to_rownames("sample_type") %>% 
    dplyr::select(where(is.numeric)) %>%
    dplyr::mutate(row = row_number()) %>% 
    drop_na()
  
  num_row_numbers = num %>% dplyr::select(row)
  
  grp = 
    dat %>% 
    dplyr::select(!where(is.numeric)) %>% 
    dplyr::mutate(row = row_number())
  #  %>% 
  #    right_join(num_row_numbers)
  
  set.seed(123456)
  jacc = vegdist(num, method = "jaccard")
  mds = metaMDS(jacc, k=3)
  
  #scores(mds)
  
  scores = 
    as.data.frame(scores(mds)) %>% 
    dplyr::mutate(row = row_number()) %>% 
    # rownames_to_column("DOC_ID") %>% 
    left_join(grp) %>% 
    left_join(cluster)
  
  
}


mds_positive = compute_nmds(dat = fticr_wide %>% filter(mode == "positive"), cluster = clusters_pos)
mds_negative = compute_nmds(dat = fticr_wide %>% filter(mode == "negative"), cluster = clusters_neg)


```


```{r icr_NMDS_plot}
library(ggConvexHull)


mds_positive %>% 
  ggplot(aes(x = NMDS1, y = NMDS2,
             color = transect_location,
             shape = region, size = horizon))+
  
  #  geom_convexhull(
  #                  aes(fill = as.character(cluster)), 
  #                color = "black",
  #                linewidth = 1.2,
  #                alpha = 0.8)+
  #  
  geom_point()+
  labs(title = "NMDS, positive mode")+
  theme(legend.position = "right")
#  scale_shape_manual(values = c(21, 22, 23))+
#  scale_fill_manual(values = pnw_palette("Bay", 5))+
# guides(fill = guide_legend(override.aes = list(shape = 21)))




mds_negative %>% 
  ggplot(aes(x = NMDS1, y = NMDS2,
             color = transect_location,
             shape = region, size = horizon))+
  
  #  geom_convexhull(
  #                  aes(fill = as.character(cluster)), 
  #                color = "black",
  #                linewidth = 1.2,
  #                alpha = 0.8)+
  
  geom_point()+
  labs(title = "NMDS, negative mode")+
  theme(legend.position = "right")

#  scale_shape_manual(values = c(21, 22, 23))+
#  scale_fill_manual(values = pnw_palette("Bay", 5))+
# guides(fill = guide_legend(override.aes = list(shape = 21)))

```


### RELATIVE ABUNDANCE

```{r icr_relabund, fig.width=12, fig.height=12}
library(ggh4x)

icr_relabundance_summary = 
  icr_relabundance %>% 
  group_by(region, transect_location, horizon, mode, Class) %>% 
  dplyr::summarise(mean = mean(relabund))

icr_relabundance_summary %>% 
  ggplot(aes(x = transect_location, y = mean, fill = Class))+
  geom_bar(stat = "identity")+
  facet_nested(mode ~ region + horizon, scales = "free_x", space = "free_x",
               nest_line = element_line(color = "grey"))
```




---


---

## Session Info 

<details>
  <summary> Session Info </summary>

Date run: `r Sys.Date()`

```{r}
sessionInfo()
```

</details>